---
title: "Viz 1"
output: github_document
date: "2024-09-26"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

Import the weather data
```{r, message=FALSE}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Basic scatter plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

saving it to a variable

```{r}
ggp_weather_scatterplot <- 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Can see that ggplot warns about missing rows:

```{r}
weather_df |>
  filter(is.na(tmax))
```

## Fancier scatter plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE)
```

Where you define aesthetics matters

This only applies color = name to the points, not to geom_smooth. geom_smooth does not 'know' about color = name.
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name),alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE)
```

use faceting
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #LHS is rows (.=nothing), RHS is columns
```

Make a more interesting scatterplot
```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(.~name)
```

Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(tmax_f = 9/5 * tmax + 32,
         tmin_f = 9/5 * tmin + 32) |>
  ggplot(aes(x = tmin_f, y = tmax_f)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

## Small things

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()
```

there is no variable "blue", so putting color = "blue" in ggplot(aes()) does not work out
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax,color = "blue")) + 
  geom_point()
```

This will work:
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(color = "blue")
```

## Univariate plots

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

How to fix this/make it better? Maybe facet?

Also not great, but better
```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() +
  facet_grid(.~name)
```

Maybe a density plot?
```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = 0.3)
```

boxplots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```

try violin plots
```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin()
```

Another option: ridge plot
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name)) + 
  geom_density_ridges()
```

Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.
```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = 0.3)
```

```{r}
weather_df |>
  ggplot(aes(x = name, y = prcp, fill = name)) + 
  geom_violin()
```

```{r}
weather_df |>
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges()
```

None of these options look to good, maybe if we removed 0 precip days?

```{r}
weather_df |>
  filter(prcp > 0) |>
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges()
```

## saving and embedding plots

```{r}
ggp_weather <- weather_df |>
  ggplot(aes(x = date, y= tmax, color = name)) +
  geom_point()

ggsave("plots/ggp_weather.pdf",ggp_weather, width = 8, height = 6)
```

Embedding plots

```{r, fig.width=6, fig.asp = 0.6}
weather_df |>
  ggplot(aes(x = date, y= tmax, color = name)) +
  geom_point()
```

