---
title: "Viz 2"
output: github_document
date: "2024-10-01"
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
```

Dr. Goldsmith uses this chunk at the top of all his docs to set the defaults
```{r, eval=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


Import the weather data
```{r, message=FALSE}
weather_df <- 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Make a fancy scatterplot: add labels
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  labs(title = "Temperature scatterplot",
       x = "Minimum temp (C)",
       y = "Maximum temp (C)",
       color = "Location",
       caption = "Weather data taken from rnoaa package for 3 stations.")
```

Scales -- start with `x` and `y` and then do `color`
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  labs(title = "Temperature scatterplot",
       x = "Minimum temp (C)",
       y = "Maximum temp (C)",
       color = "Location",
       caption = "Weather data taken from rnoaa package for 3 stations.") + 
  scale_x_continuous(breaks = c(-15,0,20),
                     labels = c("-15C","0C","20C")) + 
  scale_y_continuous(limits = c(0,30),
                     transform = "sqrt") 
```

Look at color

```{r}
ggp_scatterplot <- weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  labs(title = "Temperature scatterplot",
       x = "Minimum temp (C)",
       y = "Maximum temp (C)",
       color = "Location",
       caption = "Weather data taken from rnoaa package for 3 stations.") +
  viridis::scale_color_viridis(discrete = TRUE) 
```

## Themes

```{r}
ggp_scatterplot + 
  theme_bw() +
  theme(legend.position = "bottom")
```

Order matters here
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") +
  theme_bw() 
```

some other themes
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") +
  theme_minimal() 
```

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") +
  theme_classic() 
```

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") +
  ggthemes::theme_economist() 
```

Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot
```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point(alpha = 0.3) + 
  #geom_smooth(se = FALSE) +
  labs(title = "Max temp by date",
       x = "Date",
       y = "Max temperature (C)",
       color = "Location",
       size = "Precipitation") +
  scale_color_viridis_d() +
  theme_bw() + 
  theme(legend.position = "bottom")
```

Extra bonus stuff in `ggplot`

use different datasets in different geoms
```{r}
central_park_df <- weather_df |>
  filter(name == "CentralPark_NY")

molokai_df <- weather_df |>
  filter(name == "Molokai_HI")

molokai_df |>
  ggplot(aes(x = date, y = tmax)) + 
  geom_point() + 
  geom_line(data = central_park_df)
```

## Multiple panels
```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density() + 
  facet_grid(. ~ name)
```

combining different plots

```{r}
ggp_max_min <- weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)

ggp_tmax_density <- weather_df |>
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = 0.3)

ggp_tmax_date <- weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth()

(ggp_max_min + ggp_tmax_density) / ggp_tmax_date
```
Could add titles, combine legends, etc.

## Data manipulation

```{r}
weather_df |>
  mutate(name = fct_relevel(name, c("Molokai_HI","CentralPark_NY","Waterhole_WA"))) |>
  ggplot(aes(x = name, y = tmax, fill = name)) + 
  geom_violin(alpha = 0.5)
```

Pulse data next
```{r}
pulse_df <- haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_")  |>
  mutate(visit = if_else(visit == "bl","00m",visit))

pulse_df |>
  ggplot(aes(x = visit, y = bdi_score)) + 
  geom_boxplot()
```

pups and litters data
```{r}
pups_df <- read_csv("data/FAS_pups.csv",
                     na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  mutate(sex = case_match(sex, 
                          1 ~ "male", 
                          2 ~ "female"))

litters_df <- read_csv("data/FAS_litters.csv",
                       na = c("NA",".","")) |>
  janitor::clean_names() |>
  separate(group, into = c("treatment","day"),sep = 3)


fas_df <- inner_join(pups_df,litters_df, by = "litter_number") |>
  pivot_longer(cols = pd_ears:pd_walk,
               names_to = "outcome",
               values_to = "pd",
               names_prefix = "pd_")
```

plotting
```{r}
fas_df |>
  ggplot(aes(x = treatment,y = pd)) +
  geom_boxplot() +
  facet_grid(day ~ outcome)
```

